\plaintitle

### 2016-12-15 drafts/labor/piecework_lit.tex
notes from msb added
moved some things around and refactored the earlier parts of the piecework literature review
i didn't bother to change the narrative of the sections yet, in part because i don't really know how to rewrite it from the chronological ordering, especially going from the industrial work->farmwork sections, since taylorism was sort of a directional bridge. maybe i could write it so it's more like 'where did taylorism come from though? for that, we turn to farmwork...'?

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it; and
  \item trace the rise and fall of piecework at a very high level,
        identifying key figures and ideas during this time;
\end{inlinelist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

@@

In summary, piecework:
\begin{inlinelist}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work; and
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

@@

\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\topic{In this section we will offer a brief overview of the history of piecework;
this should not be mistaken for a comprehensive background.}
Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework in the contexts of
the early days of the Industrial Revolution,
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework.

@@

The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
Despite this, workers' needs were mostly ignored,
leading to frustration over poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives})~\cite{riisOtherSideLives}.
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries; and
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

\subsection{Case studies in piecework}
We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
\begin{numberlist}
  \item Airy's employment of \textit{human computers},
  \item domestic and farm work (in particular, the ``match--girls''), and
  \item railroad workers across the United States.
\end{numberlist}

@@

\subsubsection{Airy's Computers}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
% piecework, found in George Airy's creation of
% the British Nautical Almanac.}
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.

\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
For one thing,
Airy had set a policy of firing his computers once they reached age 23.
This practice ensured two outcomes that arguably disfavored workers.
\begin{Numberlist}
\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
\item it limited workers' ability to organize
      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
\end{Numberlist}
% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}

@@

\subsubsection{Railroad and other Industrial Workers}
\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
Scientific management argued that it was possible to measure work at
unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
As \citeauthor{Brown01041990} points out,
piecework most greatly benefits the instrumented measurement of workers,
% but certainly in Ford and Taylor's time,
but in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
as rudimentary guides on per--task compensation.
This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
railroad maintenance, automobile manufacturing, and later weapons manufacturing.
\ali{???? better word for ``weapons manufacturing''?}
% Instead, as a result,
% the distillation of work into smaller units ultimately
% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.

>>>>

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise and fall of piecework at a very high level,
        identifying key figures and ideas during this time;
\end{inlinelist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

@@

In summary, piecework:
\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work;
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

@@

\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\topic{In this section we'll discuss the history of piecework at a high level
to contextualize the observations and conclusions we later make,
rather than provide a comprehensive historical account.}
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework;
as close to a \textit{thick description} of piecework
as we can afford~\cite{geertz1973interpretation}.

@@

The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
Despite this, workers' needs were mostly ignored,
leading to frustration over poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries;
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

\subsection{Case studies in piecework}
We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
\begin{numberlist}[itemjoin*={~and~},itemjoin={~}]
  \item railroad workers across the United States;
  \item Airy's employment of \textit{human computers};
  \item domestic and farm work (in particular, the ``match--girls'').

\end{numberlist}

@@

% \ali{maybe go backwards on this}

% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}

\subsubsection{Railroad and other Industrial Workers}
\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
Scientific management argued that it was possible to measure work at
unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
As \citeauthor{Brown01041990} points out,
piecework most greatly benefits the instrumented measurement of workers,
but in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
as rudimentary guides on per--task compensation.
This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
railroad maintenance, automobile manufacturing, and later weapons manufacturing.
\ali{???? better word for ``weapons manufacturing''?}
% Instead, as a result,
% the distillation of work into smaller units ultimately
% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.

@@

\subsubsection{Airy's Computers}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
% piecework, found in George Airy's creation of
% the British Nautical Almanac.}
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
\ali{also talk about the challenge/competition Airy was in with other astronomers,
and the approach of verifying worker results}

\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
For one thing,
Airy had set a policy of firing his computers once they reached age 23.
\ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
\begin{Numberlist}
\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
\item it limited workers' ability to organize
      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
\end{Numberlist}}
